\LARGE \bf Mean Estimation from One-Bit Measurements

### 2019-09-23 problem-formulation.tex
Working on problem setting

\section{Problem Formulation \label{sec:problem}}

Let $f(x)$ be a symmetric and log-concave density function with a finite second moment $\sigma^2$. For $\theta \in \Theta$, denote by $P_X$ the probability distribution with density $f\left( x-\theta \right)$. Therefore, $P_X$ is an absolutely continuous log-concave distribution with mean $\theta$ and variance $\sigma^2$. Symmetry and log-concavity of $f(x)$ imply that $P_X$ is strongly unimodal (has a unique local maxima) with its mode at $x =\theta$ \cite{ibragimov1956composition}. We further assume that the \emph{parameter space} $\Theta$ is a closed interval of the real line. 
\par
In some situations it is useful to assume that $\theta$ is drawn once from the prior distribution $\pi$ on $\Theta$. In this case we assume that $\pi$ is an absolutely continuous distribution, and denote its density by $\pi(\theta)$, i.e., $\pi(d\theta) = \pi(\theta)d\theta$.  
\par
The random variables $X_1,\ldots,X_n$ represent $n$ independent samples from $P_X$. 
We are interested in estimating $\theta$ from a set of $n$ binary messages $B_1,\ldots,B_n$, obtained from $X_1,\ldots,X_n$ under three possible scenarios illustrated in Figure~\ref{fig:setup}: 
\begin{itemize}
\item[(i)~~] Centralized $B_i(X_1,\ldots,X_n)$, $i=1,\ldots,n$.
\item[(ii)~] Adaptive $B_i(X_i,B_1,\ldots,B_{i-1})$, $i=2,\ldots,n$.
\item[(iii)] Distributed $B_i(X_i)$, $i=1,\ldots,n$.
\end{itemize}

The performance of an estimator ${\theta}_n \triangleq {\theta}_n(B^n)$ in any of these cases is measured according to the quadratic risk:
\begin{equation}
\label{eq:error_def}
R_n \triangleq \mathbb E\left({\theta}_n - \theta \right)^2,
\end{equation}
where the expectation is taken with respect to the distribution of $X_1,\ldots,X_n$ and, whenever available, a prior distribution $\pi(\theta)$ over $\Theta$.  
%It is well known that minmax estimation error 
%\begin{equation}
%R_n = \min \sup_{\theta \in \Theta}  \mathbb E\left[ \left({\theta}_n - \theta \right)^2 | \theta \right],
%\end{equation}
%can be obtained from $R_n$ by assuming a particular least favorable prior distribution on the parameter space $\Theta$ \cite{casella1981}. 
%
The main problems we consider in this paper are the minimal value of \eqref{eq:error_def}, as a function of $n$ and $f(x)$, under different choices of the encoding functions in cases (i), (ii), and (iii). \par

We give particular attention to the ARE of estimators with respect to an asymptotically normal efficient estimator that is not subject to the bit constraint. Specifically, let $\{a_n,\,n\in \mathbb N\}$ be a sequence such that 
\[
\sqrt{a_n}\left({\theta}_n - \theta\right) \overset{d}{\longrightarrow} \Ncal(0, \sigma^2).
\]
Then the ARE of ${\theta}_n$ with respect to an unconstrained efficient estimator for $\theta$ is defined as \cite[Def. 6.6.6]{lehmann2006theory}
\[
\ARE({\theta}_n) \triangleq
\lim_{n\rightarrow \infty} \frac{a_n}{n}. 
\]
Note that in the special case where there exists $V \in \mathbb R$ such that
\[
a_n \mathbb E \left({\theta}_n - \theta \right)^2 = V + o(1),
\]
with $o(1)\to 0$ as $n\to \infty$, then the ARE of ${\theta}_n$ is finite and equals $\sigma^2/V$. \par

>>>>

\section{Problem Formulation and Notation}
\label{sec:problem}

Let $f : \R \to \R_+$ be a symmetric and log-concave probability density,
which necessarily has finite second moment $\sigma^2$, and let $\Theta
\subset \R$ be closed and convex.  For $\theta \in \R$, let $P_\theta$ be
the probability distribution with density $f(x-\theta)$, so that $\theta$
indexes the location family $\{P_\theta\}_{\theta \in \Theta}$.  The
log-concavity and symmetry $f(x)$ imply that $P_\theta$ has a unique mean
and median at $\theta$~\cite{ibragimov1956composition}.
We observe a sample $X_1, \ldots, X_n \simiid P_\theta$, where $\theta$
is unknown, and wish to estimate $\theta$ given only binary
messages $B_1, \ldots, B_n \in \{0, 1\}$ about each $X_i$.
We study this under three distinct computational scenarios, which
we illustrate in Figure~\ref{fig:setup}:
\begin{enumerate}[(i)]
\item \label{item:centralized} Centralized, where $B_i =
  B_i(X_1,\ldots,X_n)$, $i=1,\ldots,n$.
\item \label{item:adaptive} Adaptive, where $B_i =
  B_i(X_i,B_1,\ldots,B_{i-1})$, $i=2,\ldots,n$.
\item \label{item:distributed}
  Distributed, where $B_i = B_i(X_i)$, $i=1,\ldots,n$.
\end{enumerate}
\noindent
We also consider a hybrid of the fully distributed setting (where the bits
$B_i$ are independent) and the adaptive setting (where each bit $B_i$ may
depend on the previous bits) to a \emph{one-step} adaptive setting, where
the quantization scheme may be modified to depend on one fixed function of
the previous information.
\begin{enumerate}[(i')]
\setcounter{enumi}{1}
\item One-step adaptive, where for some function $g$ and
  a (fixed) $t$, if $i \le t$ then
  $B_i = B_i(X_i)$ while if $i > t$, then
  $B_i = B_i(X_i, g(B_1, \ldots, B_t))$.
\end{enumerate}

We measure the performance of an estimator $\theta_n \defeq
\theta_n(B_1^n)$ by one of a few notions. In the simplest case,
we assume a prior $\pi$ on $\theta$ (which may be a point
mass) and consider the quadratic risk
\begin{equation}
  \label{eq:error_def}
  R_n = R_n(\pi) \defeq \int \E_\theta\left({\theta}_n - \theta \right)^2
  d\pi(\theta),
\end{equation}
where the expectation is taken with respect to the distribution of
$X_1,\ldots,X_n \simiid P_\theta$.  The main problems we consider in this
paper are the minimal value of the risk~\eqref{eq:error_def} as a function
of the sample size $n$ and the density $f$, under different choices of the
encoding functions in
cases~\eqref{item:centralized}--\eqref{item:distributed}.
The quadratic risk~\eqref{eq:error_def} may be infinite in some cases;
we defer discussion of this case to later sections, as it is technically
demanding and detracts from the presentation here.

Now, let $\sigma_f^2 \defeq \E[\frac{f'(X)^2}{f(X)^2}]$ be the Fisher
information for the location in the family $\{P_\theta\}$, which is finite
when $f$ is log-concave and symmetric. We give particular attention to the
asymptotic relative efficiency (ARE) of estimators with respect to
asymptotically normal efficient estimators achieving the information
bound~\cite{VanDerVaart98}. In this case,
if $\{m(n), n \in \N\}$ is a sequence such that
\begin{equation*}
  \sqrt{m(n)} (\theta_n - \theta) \cd \normal(0, \sigma_f^2),
\end{equation*}
then the ARE of the estimator is~\cite[Def.~6.6.6]{LehmannCa98}
\begin{equation*}
  \ARE({\theta}_n) \defeq
  \liminf_{n\rightarrow \infty} \frac{m(n)}{n}.
\end{equation*}
In the special case where there exists $V \in \mathbb R$ such that
\[
m(n) R_n =
m(n) \mathbb E_\theta\left({\theta}_n - \theta \right)^2 = V + o(1),
\]
the ARE of ${\theta}_n$ is $\sigma_f^2/V$, so that $\theta_n$ requires a
sample $V / \sigma_f^2$-times larger than that of an efficient estimator for
comparable accuracy to the (information) efficient estimator.

%%% JCD: I AM HERE BUT STILL GOING THROUGH THINGS %%%
